Risk Perceptions for Wearable Devices

### 2015-03-04 tex-inputs/intro.tex
started to look at results
intro redo

Because interactions with the wearables so frictionless, they are a popular choice for augmenting life with novel feedback and taking advantage of pre-existing applications more seamlessly. The constantly captured data from these devices has many benefits, ranging form a more natural, human-centered interface experience to healthier, fitness-data inspired lifestyle. It is clear why wearable devices are becoming even more popular, especially as they have more capabilities and benefits.

Despite their popularity, the market is still in its infancy. It is hard to predict what device and applications will be most popular. Therefore, we use the current interest as a baseline. The most common wearables at the time of this study were in the form of fitness trackers and smart watches, with a survey of 3,956 wearables-enthusiastic respondents reporting the most popular devices as fitness bands (61\%), smart watches (45\%) and other health devices (17\%)~\cite{Nilsen}.  However, there are also wearable glasses \cite{ 2_google_2014, 3_sony_global_2014}, rings \cite{4_ringly_2014}, bracelets \cite{5_intel_2014} and more. New wearables trends for 2015 include smart clothes made out fiber-optic fabrics which give information about workouts and UV technology hairslide which measures sun exposure \cite{1_digital_trends_2014, 2_arthur_2014}.

Wearable devices bring with them new potential security and privacy concerns. Many concerns expose users' activities without their awareness or consent. For instance, Fitbit's fitness profiles were public by default and also allowed sex to be tracked as exercise~\cite{Fitbit}, resulting in the inadvertent disclosure of sensitive information. Additionally, public discomfort prevented Google from enabling capabilities such as facial detection on Google Glass \cite{GlassDetection}.

Although Google Glass was most iconic wearable of its time, it has since disappeared due to complaints from the public\cite{13_google_2015}. No formal statement from Google addresses the states that the reason for its disappearance is due to negative social pressures. However, most suspect that this is the reason for its disappearance, rather than a shift in interest. One popular tech magazine called glass the ``straw the broke the back of the privacy camel's back \cite{14_dvorak_2014},'' and made bystanders feel uncomfortable with the possibility of recording personal details, conversations, and photos. Many instances of assault \cite{1_russell_2014, 15_mashable_2014, 16_gross_2014} on wearers were reported, resulting from a combination of privacy concerns and social tension\footnote{Google has come to represent gentrification in San Francisco ... [which] makes wearable device owners look and sound like one of the people whom residents of the city have come to feel oppressed by \cite{17_matyszczyk_2014}. The term ``Glasshole'' \cite{18_google_2015} has been defined specifically to refer to Google glass wearers . The public was generally unsympathetic to the assults aggravated by social tension. Tweeting and comments included: ``I'm going to shove your pathetic nerd body into a locker,'' ``Nothing radicalizes you like having [your] home taken from you.'' and``As far as your 'hate crime' allegation is concerned, Entitled Tech Trash isn't a protected class just yet'' \cite{1_russell_2014,15_mashable_2014}.}.

{\color {red} This calls for privacy research in this area! It's more important than ever! Or something along those lines.}  To avoid scandalous breaches of privacy and public opposition to new capabilities, it is critical we understand user concerns surrounding wearable devices to prevent negative social impact and breaches of privacy. There has been extensive work on the need for showing users and bystanders how certain data is being captured so that they can have more control. We have an issue of showing everyone everything all the time, since attention is a finite resources. We need to identify situations that people want to warn them about, and transparently access things that are benign. A better understanding of users' risk perceptions will enable researchers and companies to focus on users' concerns. The goal of this paper is to gain a better sense of what those concerns are.

>>>>

The constantly captured data from these devices has many benefits, ranging from a more natural, human-centered interface experience to a healthier, fitness-data inspired lifestyle. It is clear why wearable devices are becoming even more popular, especially as they have more capabilities and benefits over traditional devices.

Along with these benefits, wearable devices bring with them new potential privacy issues which expose users' activities without their awareness or consent. Fitbit allowed sex to be tracked as exercise while fitness profiles were public by default~\cite{Fitbit}, resulting in the inadvertent disclosure of sensitive information. Public discomfort toward facial recognition was so strong that the capability was not even deployed for Google Glass \cite{GlassDetection}. Google Glass, an iconic wearable of today, has since disappeared \cite{13_google_2015}. Most suspect that the reason for its disappearance was because it was the ``straw the broke the back of the privacy camel's back'' \cite{14_dvorak_2014}. Some Glass wearers faced assault \cite{1_russell_2014, 15_mashable_2014, 16_gross_2014}, while bystanders felt uncomfortable personal details, conversations, and photos possibly recorded.

{\color {red} We should warn people about such things. But we can't do that, since attention is a finite resource. This calls for privacy research in this area!}  To avoid scandalous breaches of privacy and public opposition to new capabilities, it is critical we understand user concerns surrounding wearable devices to prevent negative social impact and breaches of privacy. There has been extensive work on the need for showing users and bystanders how certain data is being captured so that they can have more control. We have an issue of showing everyone everything all the time, since attention is a finite resource. We need to identify situations that people want to warn them about, and transparently access things that are benign. A better understanding of users' risk perceptions will enable researchers and companies to focus on users' concerns. The goal of this paper is to gain a better sense of what those concerns are.

{\color {red} smart phone concerns. we haven't gotten it right for this yet! We need to do something for wearables before it's too late.}
